Multilevel generalizations of collective spin dynamics and squeezing

### 2019-10-23 collective_sun/multilevel_spin_notes.tex
cleaned up the derivation of an effective weak-SOC hamiltonian

\DeclareMathOperator{\sign}{sign}

@@

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
of operators that permute tensor factors of $\H_n^{\otimes N}$
according to elements $\sigma\in\SS_N$ of the symmetric group of order
$N$.  We then expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue $1$ with respect to all pairwise permutation operators
$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
symmetric manifold $\M_\FS\subset\H_n^{\otimes N}$ is spanned by a
basis $\set{\ket{\v m}}$ of fully symmetric states that can be
uniquely identified by a list of integers
$\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where $m_s$
indicates the total occupation number of subsystem state $s$, and
$\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$, and energy of $\ket{\psi_\FS}\in\M_\FS$ with
respect to the interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int_perm} is $-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.

In addition to the fact that the fully symmetric manifold $\M_\FS$ is
ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
important feature of $\M_\FS$ is that it is closed under the action of
unitaries generated by collective operators of the form
$\Q=\sum_pQ^{(p)}$, as well as linear combinations and products
thereof.  Closure of $\M_\FS$ under collective dynamics is a
straightforward consequence of the permutational symmetry obeyed by
collective operators.

@@

When SOC is weak,

In the regime of weak SOC with $\abs{\phi}\ll1$, by considering
\begin{enumerate*}
\item all $L\le12$ for arbitrary $n$, and
\item all $L<n$ for all $n\le25$
\end{enumerate*}
(and conjecturing the same functional form for all $L$ and $n$), we
find that
% expand Clebsch-Gordan coefficients in a power series according to
% \url{functions.wolfram.com/07.38.06.0008.01}
\begin{align}
  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
  \label{eq:A_L_phi_small}
\end{align}
where
\begin{align}
  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
  \sqrt{2L+1}\, \f{L!}{\p{2L+1}!},
  &&
  n_L \equiv \prod_{\ell=-L}^L\p{n-\ell}.
\end{align}

@@

>>>>

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\cov}{cov}
\let\var\relax
\DeclareMathOperator{\var}{var}

@@

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
of operators that permute tensor factors of $\H_n^{\otimes N}$
according to elements $\sigma\in\SS_N$ of the symmetric group of order
$N$.  We then expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue $1$ with respect to all pairwise permutation operators
$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
symmetric manifold $\M_\FS$ is spanned by a basis $\set{\ket{\v m}}$
of fully symmetric states that can be uniquely identified by a list of
integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where
$m_s$ indicates the total occupation number of subsystem state $s$,
and $\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$, and the energy of a state
$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
$-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.

In addition to the fact that the fully symmetric manifold $\M_\FS$ is
ground-state manifold of the interaction Hamltonian $H_{\t{int}}$, an
important feature of $\M_\FS$ is its closure under the action of
unitaries generated by collective operators of the form
$\Q=\sum_pQ^{(p)}$, as well as linear combinations and products
thereof.  Closure of $\M_\FS$ under collective dynamics is a
straightforward consequence of the permutational symmetry obeyed by
collective operators.

@@

The single-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ breaks
permutational symmetry, thereby coupling states within the fully
symmetric manifold $\M_\FS$ to states outsides it.  If SOC is
sufficiently weak, however, transitions outside the fully symmetric
manifold $\M_\FS$ are energetically suppressed by interactions, and we
can account for the effect of SOC perturbatively.  In the regime of
weak SOC with $\abs{\phi}\ll1$, by considering
\begin{enumerate*}
\item all $L\le12$ for arbitrary $n$, and
\item all $L<n$ for all $n\le25$
\end{enumerate*}
(and conjecturing the same functional form for all $L$ and $n$), we
find that
% expand Clebsch-Gordan coefficients in a power series according to
% \url{functions.wolfram.com/07.38.06.0008.01}
\begin{align}
  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
  &&
  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
  \f{\sqrt{2L+1}\,L!}{\p{2L+1}!},
  &&
  n_L \equiv \prod_{\ell=-L}^L\p{n+\ell}.
  \label{eq:A_L_phi_small}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Weak SOC in the fully symmetric manifold}

Here we derive the effective Hamiltonian induced on the fully
symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
spacing $a=1$ throughout this section.  Perturbations $V$ to the
interaction Hamiltonian $H_{\t{int}}$ induce the following effective
Hamiltonians on the fully symmetric manifold
$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  &&
  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
  &&
  \P_0 = \sum_{E_{\varepsilon}=0} \op{E_{\varepsilon}},
  &&
  \E \equiv \sum_{E_\varepsilon>0} \f{\op{E_\varepsilon}}{E_\varepsilon},
  \label{eq:pert}
\end{align}
where $\ket{E_\varepsilon}$ is an eigenstate of the interaction
Hamiltonian $H_{\t{int}}$ with energy $E_\varepsilon$ relative to the
interaction energy of fully symmetric states
$\ket{\psi_\FS}\in\M_\FS$.  The sum over all $E_\varepsilon>0$ is a
sum over a complete basis of states $\set{\ket{E_\varepsilon}}$ for
the orthogonal complement of $\M_\FS$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Effective SOC at first order}
\label{sec:H_eff_1}

The single-particle SOC Hamiltonian is
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} D_{L,0}^{(q)},
  &&
  w_L\p{q} \equiv
  \begin{cases}
    \cos q & L~\t{even} \\
    \sin q & L~\t{odd}
  \end{cases},
  \label{eq:H_lat_SOC_A}
\end{align}
where $A_L^{(\phi)}$ is essentially a sum over Clebsch-Gordan
coefficients, defined in \eqref{eq:A_L_phi}, and scales as
$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ when $\abs{\phi}\ll1$.
When the SOC angle $\phi$ is sufficiently small in magnitude, the
effective Hamiltonian induced on the fully symmetric manifold $\M_\FS$
at first order in the perturbation $H_{\t{lat}}^{(\phi)}$ is
\begin{align}
  H_{\t{eff}}^{(1)}
  = \P_0 H_{\t{lat}}^{(\phi)} \P_0
  = -t \sum_{q,L} w_L\p{q} A_L^{(\phi)} \P_0 D_{L,0}^{(q)} \P_0,
\end{align}
where we can freely change the index of each single-mode drive
operator $D_{L,0}^{(q)}$ within the fully symmetric manifold to
re-write
\begin{align}
  \P_0 D_{L,0}^{(q)} \P_0 = \f1N \sum_p \P_0 D_{L,0}^{(p)} \P_0
  = \f1N \P_0 \D_{L,0} \P_0,
\end{align}
which implies that
\begin{align}
  H_{\t{eff}}^{(1)}
  = -t \sum_L \EE_q\sp{w_L\p{q}} A_L^{(\phi)} \D_{L,0},
  &&
  \EE_q\sp{X_q} \equiv \f1N \sum_q X_q.
  \label{eq:H_eff_1}
\end{align}
Here $\EE_q\sp{X_q}$ is an average of $X_q$ over all occupied
quasi-momenta $q$.  Note that we have dropped the explicit projection
operators $\P_0$ in \eqref{eq:H_eff_1} with the understanding that we
are implicitly restricting all effective dynamics to the fully
symmetric manifold $\M_\FS$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Effective SOC at second order}
\label{sec:H_eff_2}

In order to expand the second-order effective Hamiltonian
$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
\eqref{eq:H_lat_SOC_A} when the SOC angle $\abs{\phi}\ll1$, we note
that $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} takes the form
\begin{align}
  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
  \label{eq:w_chi}
\end{align}
for some appropriately defined coefficients $\chi_{Ls}$, which implies
that
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
  &&
  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(2)}
  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
  \label{eq:H_eff_2_start}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold
$\M_\FS$, and $\E$ is a sum over excited-state projectors with
energetic suppression factors, defined in \eqref{eq:pert}.  The
remainder of this section is devoted to the task of simplifying
\eqref{eq:H_eff_2_start} to arrive at our final result in
\eqref{eq:H_eff_2}, expressed in purely in terms of collective drive
operators.  We will break up this task into three steps:
\begin{enumerate*}
\item eliminating the excited-state projectors in $\E$,
\item evaluating the product of operators in \eqref{eq:H_eff_2_start},
  and
\item evaluating the sum over $r,s\in\set{\pm1}$.
\end{enumerate*}

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Mean-subtracted spin-wave operators and the elimination
  of excited-state projectors}

In order to simplify the product of operators in
\eqref{eq:H_eff_2_start}, we note that sandwiching the ``spin-wave''
operators $\Z_{Jr},\Z_{Ks}$ by the ground-state projector $\P_0$ and
the excited-state projectors in $\E$, picks off the components of
these spin wave operators that are off-diagonal with respect to the
symmetric manifold $\M_\FS$ and its orthogonal complement.  We
therefore define the mean-subtracted spin-wave operators
  \tilde\Z_{Ks}
  \equiv \sum_q \p{e^{isq} - \EE_k\sp{e^{isk}}} D_{K,0}^{(q)},
  &&
  \EE_k\sp{X\p{k}} \equiv \f1N \sum_k X\p{k},
for which the difference $\tilde\Z_{Ks}-\Z_{Ks}\propto\D_{K,0}$ is a
collective operator that only couples fully symmetric states
$\ket{\psi_\FS}\in\M_\FS$ to fully symmetric states.  In particular,
  \P_0 \tilde\Z_{Jr} \E \tilde\Z_{Ks} \P_0
  = \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
which allows us to freely take
$\Z_{Jr},\Z_{Ks}\to\tilde\Z_{Jr},\tilde\Z_{Ks}$ with no consequence
for the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:H_eff_2_start}.  Mean-subtracted spin-wave operators
$\tilde\Z_{Ks}$ have the convenient feature that their action on any
fully symmetric state $\ket{\psi_{\FS}}\in\M_\FS$ yields an
eigenvector of the interaction Hamiltonian
$H_{\t{int}}=-u\sum_{p<q}\Pi_{pq}$.  To substantiate this claim, we
expand
  \sum_{p<q} \Pi_{pq} \tilde\Z_{Ks} \ket{\psi_\FS}
  = \sum_{\substack{p<q\\k}}
  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}.
  \label{eq:Z_eig_start}
If $k\notin\set{p,q}$, then the permutation operator $\Pi_{pq}$
commutes with $D_{K,0}^{(k)}$, and in turn acts trivially on the
permutationally-symmetric state $\ket{\psi_\FS}$, giving us
  \sum_{\substack{p<q\\k\notin\set{p,q}}}
  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  = { N-1 \choose 2 } \tilde\Z_{Ks} \ket{\psi_\FS}.
  \label{eq:Z_eig_notin}
If $k\in\set{p,q}$, then we can resolve the identity operator
$\1=\Pi_{pq}^\dag \Pi_{pq}$ to expand
\begin{multline}
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  = \f12 \sum_{\substack{p\ne q\\k\in\set{p,q}}}
  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  \Pi_{pq} D_{K,0}^{(k)} \Pi_{pq}^\dag \Pi_{pq} \ket{\psi_\FS} \\
  = \sum_q \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  D_{K,0}^{(q)} \ket{\psi_\FS},
\end{multline}
where $\sum_k \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}} = 0$, so
  \sum_{k\ne q} \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  = - \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}},
and in turn
  \sum_{\substack{p<q\\k\in\set{p,q}}}
  \p{e^{isk} - \EE_\ell\sp{e^{is\ell}}}
  \Pi_{pq} D_{K,0}^{(k)} \ket{\psi_\FS}
  = -\sum_q \p{e^{isq} - \EE_\ell\sp{e^{is\ell}}}
  D_{K,0}^{(q)} \ket{\psi_\FS}
  = - \tilde\Z_{Ks} \ket{\psi_\FS}.
  \label{eq:Z_eig_in}
Substituting the results in \eqref{eq:Z_eig_in} and
\eqref{eq:Z_eig_notin} into \eqref{eq:Z_eig_start} gives us
  \sum_{p<q} \Pi_{pq} \tilde\Z_{Ks} \ket{\psi_\FS}
  = \sp{{ N-2 \choose 2 } - 1} \tilde\Z_{Ks} \ket{\psi_\FS}.
The image of any fully symmetric state $\ket{\psi_\FS}\in\M_\FS$ under
the action of the mean-subtracted spin-wave operator $\tilde\Z_{Ks}$
is thus proportional to an eigenstate of $H_{\t{int}}$ with
interaction energy $-u\sp{{ N-1 \choose 2 } - 1}$, which is $Nu$ above
the interaction energy of $-u{N\choose2}$ of states in the fully
symmetric manifold, so
  \E \tilde\Z_{Ks} \P_0
  =  \f1{Nu} \tilde\Z_{Ks} \P_0
  = \f1{fU} \tilde\Z_{Ks} \P_0,
  \label{eq:proj_elim}
where $f\equiv N/L$ is the filling fraction of quasi-momentum modes.
We have thus far simplified the effective Hamiltonian in
\eqref{eq:H_eff_2_start} to
  H_{\t{eff}}^{(2)}
  = -\f{t^2}{4fU} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0,
  \label{eq:H_eff_2_mid}
leaving us the task of evaluating the product
$\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$, as well as the sum over
$r,s$.

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{The spin-wave operator product}

To evaluate the product $\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$ of
mean-subtracted spin-wave operators projected by $\P_0$ onto the fully
symmetric manifold $\M_\FS$, we expand
\begin{align}
  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
  = \sum_{p,q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0.
  \label{eq:ZZ_prod_start}
\end{align}
Using the permutational symmetry of the fully symmetric manifold, we
can simplify the $p=q$ terms above to
\begin{align}
  \sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
  \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}}
  \P_0 D_{J,0}^{(p)} D_{K,0}^{(p)} \P_0
  = N \cov_{p,q}\sp{e^{irp},e^{isq}}
  \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0,
\end{align}
where
\begin{align}
  \cov_{p,q}\sp{X_p,Y_q}
  \equiv \EE_p\sp{\p{X_p-\EE_k\sp{X_k}}\p{Y_p-\EE_\ell\sp{Y_\ell}}}
  = \EE_k\sp{X_k Y_k} - \EE_p\sp{X_p} \EE_q\sp{X_q}
  \label{eq:cov}
\end{align}
is the covariance between $X_p$ and $Y_q$, evaluated over all occupied
quasi-momenta $p,q$.  Similarly, the $p\ne q$ terms in
\eqref{eq:ZZ_prod_start} simplify to
\begin{multline}
  \sum_{p\ne q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
  = \sum_p \p{e^{irp}-\EE_k\sp{e^{irk}}}
  \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
  \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
\end{multline}
where $\sum_q \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}} = 0$, so
\begin{align}
  \sum_{q\ne p} \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
  = - \p{e^{isp}-\EE_\ell\sp{e^{is\ell}}},
\end{align}
and in turn
\begin{align}
  \sum_{p\ne q} \p{e^{irp}-\EE_k\sp{e^{irk}}}
  \p{e^{isq}-\EE_\ell\sp{e^{is\ell}}}
  \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0
  = - N \cov_{p,q}\sp{e^{irp},e^{isq}}
  \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0.
\end{align}
Thus far, we have that
\begin{align}
  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
  = N \cov_{p,q}\sp{e^{irp},e^{isq}}
  \p{\P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
    - \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0}.
  \label{eq:ZZ_prod_mid}
\end{align}
In order to write this result in terms of collective operators, we
expand
\begin{multline}
  \P_0 \D_{J,0} \D_{K,0} \P_0
  = \sum_p \P_0 D_{J,0}^{(p)} D_{K,0}^{(p)} \P_0
  + \sum_{p\ne q} \P_0 D_{J,0}^{(p)} D_{K,0}^{(q)} \P_0 \\
  = N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  + N \p{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0,
\end{multline}
which implies that
\begin{align}
  - N \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0
  = \f{N}{N-1} \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0.
\end{align}
The last ingredient we need to re-write \eqref{eq:ZZ_prod_mid} in
terms of collective operators is
\begin{align}
  N \P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
  = N \sum_L f_{J,0;K,0}^{L,0} \P_0 D_{L,0}^{(0)} \P_0
  = \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0,
\end{align}
where $f_{J,0;K,0}^{L,0}$ is a structure constant of the transition
operator algebra, defined in \eqref{eq:trans_prod} of Appendix
\ref{sec:trans_prod}.  Altogether, we then have that
\begin{align}
  N\p{\P_0 D_{J,0}^{(0)} D_{K,0}^{(0)} \P_0
    - \P_0 D_{J,0}^{(0)} D_{K,0}^{(1)} \P_0}
  = \f{N}{N-1} \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0
  - \f1{N-1} \P_0 \D_{J,0} \D_{K,0} \P_0,
  \label{eq:eff_sym_drive}
\end{align}
which, substituted into \eqref{eq:ZZ_prod_mid}, implies that
\begin{align}
  \P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0
  = -\f{\cov_{p,q}\sp{e^{irp},e^{isq}}}{N-1}
  \p{\P_0 \D_{J,0} \D_{K,0} \P_0
    - N \sum_L f_{J,0;K,0}^{L,0} \P_0 \D_{L,0} \P_0}.
  \label{eq:ZZ_prod}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Putting it all together}

Having evaluated the product $\P_0 \tilde\Z_{Jr} \tilde\Z_{Ks} \P_0$
of spin-wave operators in \eqref{eq:ZZ_prod}, the last ingredient
needed to simplify the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:H_eff_2_mid} is an evaluation of the sum over
$r,s\in\set{\pm1}$.  We can use the implicit definition of the
coefficients $\chi_{Ls}$ in \eqref{eq:w_chi} together with the
bilinearity of the covariance, defined in \eqref{eq:cov}, to simplify
\begin{align}
  \f14 \sum_{r,s\in\set{\pm1}} \chi_{Jr} \chi_{Ks}
  \cov_{p,q}\sp{e^{irp},e^{isq}}
  = \cov_{p,q}\sp{w_J\p{p},w_K\p{q}}.
  \label{eq:sign_sum}
\end{align}
Substituting the operator product in \eqref{eq:ZZ_prod} into the
effective Hamiltonian $H_{\t{eff}}^{(2)}$ in \eqref{eq:H_eff_2_mid}
and evaluating the sum over $r,s\in\set{\pm1}$ according to
\eqref{eq:sign_sum} gives us
\begin{align}
  H_{\t{eff}}^{(2)}
  = \f{t^2}{\p{N-1}fU} \sum_{J,K} \cov_{p,q}\sp{w_J\p{pa},w_K\p{qa}}
  A_J^{(\phi)} A_K^{(\phi)}
  \p{\D_{J,0} \D_{K,0} - N \sum_L f_{J,0;K,0}^{L,0} \D_{L,0}},
  \label{eq:H_eff_2}
\end{align}
where we have recovered factors of the lattice spacing $a$ as
appropriate, and we have dropped the explicit projection operators
$\P_0$ with the understanding that we are implicitly restricting all
effective dynamics to the fully symmetric manifold $\M_\FS$.  Note
that as the spatial filling fraction $f\equiv N/L\to1$, such that sums
over occupied quasi-momenta $q$ become sums over nearly all angles
$q\in\ZZ_N\times2\pi/N$, when both $J,K$ are even or odd
(i.e.~$J=K~\t{mod}~2$) the covariance in \eqref{eq:H_eff_2} becomes
$\cov_{p,q}\sp{\sin\p{p},\sin\p{p}}\to1/2$ or
$\cov_{p,q}\sp{\cos\p{p},\cos\p{p}}\to1/2$, whereas when one of $J,K$
is even and the other is odd (i.e.~$J\ne K~\t{mod}~2$), this
covariance becomes $\cov_{p,q}\sp{\cos\p{p},\sin\p{q}}\to0$.  In the
case of SU(2), the effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:H_eff_2} reduces to the one-axis twisting Hamiltonian
derived in Ref.~\cite{he2019engineering}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extracting dependence on the SOC angle $\phi$}

The effective Hamiltonians in \eqref{eq:H_eff_1} and
\eqref{eq:H_eff_2} are respectively first and second order in the SOC
Hamiltonian $H_{\t{lat}}^{(\phi)}$.  Here we extract the leading and
next-to-leading order dependence of these effective Hamiltonians on
the SOC angle $\phi$.  Mindful of the expansion for the coefficients
$A_L^{(\phi)} \sim \phi^L + O\p{\phi^{L+2}}$ in
\eqref{eq:A_L_phi_small}, we define the re-scaled collective drive
operators
\begin{align}
  \tilde\D_{LM} \equiv c_L \sqrt{n_L} \D_{LM},
\end{align}
in terms of which
\begin{align}
  H_{\t{eff}}^{(1)}
  &= -\phi t \p{\EE_p\sp{\sin\p{pa}} \tilde\D_{1,0}
    + \EE_q\sp{\cos\p{qa}} \tilde\D_{2,0}}
  + O\p{\phi^3},
  \label{eq:H_eff_1_phi} \\
  H_{\t{eff}}^{(2)}
  &= \f{\p{\phi t}^2}{\p{N-1}fU} \var_q\sp{\sin\p{qa}}
  \p{\tilde\D_{1,0}^2 + 2 N \tilde\D_{2,0}}
  + O\p{\phi^3},
  \label{eq:H_eff_2_phi}
\end{align}
where
\begin{align}
  \var_q\sp{X_q} \equiv \cov_{p,q}\sp{X_p,X_q}
  = \EE_p\sp{X_p^2} - \EE_q\sp{X_q}^2
\end{align}
is the variance of $X_q$ over all occupied quasi-momenta $q$, and we
have neglected scalar terms $\sim\D_{0,0}\propto\1$ that have no
physical consequence.  The effective Hamiltonians in
\eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi} reduce to those
derived in Ref.~\cite{he2019engineering} in the case of SU(2).
